Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-10 old/culled.tex
Updates from ShareLaTeX

 \documentclass[conference]{IEEEtran}
\usepackage{subfig}
\usepackage{wrapfig}
 \usepackage{amsmath}
 \usepackage{url}
 \usepackage{pifont}
 \usepackage{times}
\usepackage{rotating}
%\usepackage{balance} 
\usepackage{color, colortbl}
\usepackage{graphicx}
\usepackage{algorithmicx}
\usepackage{program}
\usepackage{cite}
\usepackage{alltt}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sec:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\definecolor{lightgray}{gray}{0.975}
\usepackage{fancyvrb}
\usepackage{stfloats}
\usepackage{multirow}
\usepackage{listings}
\usepackage{amsmath} 
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}

\definecolor{darkgreen}{rgb}{0,0.3,0}

\usepackage[table]{xcolor}
\definecolor{Gray}{rgb}{0.88,1,1}
\definecolor{Gray}{gray}{0.85}
\definecolor{Blue}{RGB}{0,29,193}
\newcommand{\G}{\cellcolor{green}}
\newcommand{\Y}{\cellcolor{yellow}}

\definecolor{MyDarkBlue}{rgb}{0,0.08,0.45} 
\newenvironment{changed}{\par\color{MyDarkBlue}}{\par}

\newcommand{\ADD}[1]{\textcolor{MyDarkBlue}{{\bf #1}}}
\newcommand{\addit}[1]{\begin{changed}\input{#1}\end{changed}}

\usepackage{color}
\usepackage{listings}
\usepackage{setspace}

\definecolor{Gray}{gray}{0.9}
\newcommand{\kw}[1]{\textit{#1}}
\newcommand{\quart}[4]{\begin{picture}(75,6)
  {\color{black}\put(#3,3){\circle*{2.5}}\put(#1,3){\line(1,0){#2}}}\end{picture}}
% New Commands

\definecolor{Code}{rgb}{0,0,0}
\definecolor{Decorators}{rgb}{0.5,0.5,0.5}
\definecolor{Numbers}{rgb}{0.5,0,0}
\definecolor{MatchingBrackets}{rgb}{0.25,0.5,0.5}
\definecolor{Keywords}{rgb}{0,0,1}
\definecolor{self}{rgb}{0,0,0}
\definecolor{Strings}{rgb}{0,0.63,0}
\definecolor{Comments}{rgb}{0,0.63,1}
\definecolor{Backquotes}{rgb}{0,0,0}
\definecolor{Classname}{rgb}{0,0,0}
\definecolor{FunctionName}{rgb}{0,0,0}
\definecolor{Operators}{rgb}{0,0,0}
\definecolor{Background}{rgb}{1,1,1}

\lstnewenvironment{python}[1][]{
\lstset{
mathescape,
numbers=#1,
numberstyle=\footnotesize,
numbersep=0.5em,
xleftmargin=0em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\small\setstretch{0.8},
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap, project, furthest, split, WHERE,clusterer, getContours, envied, fWeight, nearestContour, projection, mutate, HERE, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
%
}}{}

\title{Learning Actionable Analytics 
 (with applications for reducing defects and reducing runtimes)}
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
Rahul Krishna, Tim Menzies, Xipeng Shen\\
       \affaddr{Computer Science, NcState, USA}\\
       \{i.m.ralk, tim.menzies, xipengshen\}@gmail.com
% 3rd. author
 % use '\and' if you need 'another row' of author names
% 4th. author
\and
 Andrian Marcus\\
       \affaddr{Computer Science, UtDallas, USA}\\
       {amarcus@utdallas.edu} }

  \pagestyle{plain}
\begin{document}
  \maketitle

  % make the title area

  \begin{abstract}
 As business users demand more insightful
 analytics, data scientists need to change
 their tools. Instead of merely predicting 
 some result, they also need tools that generate ``policy advice'';
 i.e. suggestions on  how to change values in order to
 improve on some predicted value.

 This paper proposes one such policy
 generator. ``HOW'' is a near-linear-time
 tool for proposing changes to independent
 variables in order to improve on 
 the predictions of the dependent variables. Unlike other approaches
 for learning optimizations to software does not require
 an underlying of the domain (rather, HOW automatically
 learns that model as part of its operation).

 This paper uses  HOW to learn methods
 to reduce defects and decrease program runtime.
 For the test data of this paper, the improvements generated by HERE can reduce
 defect counts as well as runtimes by up to
 50\% of their original values.
  \end{abstract}
  \begin{IEEEkeywords}
Instance-based reasoning, defect prediction, random forest, optimization.
  \end{IEEEkeywords}

\section{Introduction}
Business  users are now demanding that data mining tools
be augmented with tools to support  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
that follow the mining, rather than just  the mining. For example:
\bi
\item Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
In these two points, we say that a ``plan'' is a  set of proposed changes in order to better achieve some goal. 
Generating {\em plans} is a  very different task the standard
data mining tasks of {\em predicting} some discrete or numeric class.
Predictors
and classifiers only tell us what {\em is} while planners 
  tell us {\em what to change}. For the kind of business user
discussed above, planners are the preferred to mere prediction systems
since, as one gruff user once shouted at us:
\begin{quote}
{\em ``Don't tell me what \underline{is};
tell me what to \underline{change}.''}
\end{quote}
In order to satisfy the demands of this new kind of business user, we  study the generation of {\em plans} in the context
of software analytics.  
This paper presents
HOW, a novel  {\em instance-based planner} that runs in near-linear time.
In the experiments shown below, we see that HOW's plans can
significantly reduce the expected
value of:
\bi
\item Software defects in certain object-oriented JAVA systems listed in \tion{oo}:  Ant Camel, Ivy, Jedit, JBeans, Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces;
\item The runtime of the software systems discussed in \tion{cpu}: Berkeley DB, Apache, SQLite, LLVM, 
  x264.
\ei
HOW's instance-based reasoning  makes its conclusions by
studying the space of examples. This is a very different approach to
{\em model-based planners} that use the model
as sub-routine within a Monte Carlo simulation~\cite{me07f} or the evolutionary 
programming methods explored by the search-based software
engineering community~\cite{krall14,harman12dec}. 
We find   HOW's instance-based approach useful since it can be applied
to domains that lack a model, but possess a historical log of observed behaviours.

Instance-based reasoning is also useful if building and certifying a model   takes a long time. Such slow-to-build models exist in software engineering. For  example, in previous work~\cite{me07f} we used models
developed by Boehm's group at the University of Southern California.
Some of those models took decades to develop and mature (from 1981~\cite{boehm81} to 2000~\cite{boehm00b}).  This is not an isolated example. One of us (Menzies) teaches a graduate class in automated model-based software engineering. 
Students in that class are asked to find (or build) a non-trivial model of
some software process (where ``non-trivial'' means something that processes
a  model at least an order of magnitude more complex than Brooks' Law). 
While the  number of such models has been growing in recent years\footnote{See the
catalogue at goo.gl/nv2AVK.}, 
the total number remains very small-- which is an observation consistent with
SE models typically require  much effort to build and certify.

Finally, even when there is an existing model, instance-based reasoning can be useful.
Sometimes  it is easier to access instances of a model's behaviour than the model
itself. For example, in prior work with Martin  Feather from the Jet Propulsion 
Laboratory~\cite{fea02a},  our research partner could not share a
propriety model from within the NASA firewalls. However, they could share 
logs of the input/output of that model.

In summary, this paper makes five contributions:
\be
\item A general framework for augmenting a prediction system such that
it becomes a planning system;
\item A  test that recognizes  when we should {\em not}  plan: in summary, our kind of planning is
not recommended in domains that  generate poorly performing predictors;
\item A general verification rig for assessing the value of plans generated within this rig.
\item The HOW  instance-based planner.
\item A proof that HOW runs in near-linear time.
\item An experimental validation that the HOW planner is useful for reducing the
expected values of software defects and software runtimes (at least, for the data sets studied here).
\ee
The rest of this paepr is structued as follows.

\section{From ``Prediction'' to ``Planning''}
A major  problem with building a planning mechanism in software engineering is ``how to verify the generated
plans?''. For example,  in the experiments shown below,  HOW comments  on how to reduce
software defects by adjusting certain static code parameters of the code\footnote{Such parameters can be automatically adjusted using code refactoring
tools, as shown by Binkely et al. in 2006~\cite{Binkley2006} and more recently by Mkaouer et al. in 2015~\cite{Mkaouer15}.}. How are we to verify that HOW's recommendations actually work?

Some organizations have the resources to 
run repeated trials to assess the effect of different treatments.
For example, in one remarkable recent study, Bente et al. report results
were the same specification was developed into a product by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
to (say) take a code base, assign it to $N$ different teams, apply
HOW's plans the code used by  $N/2$ teams, then check in 12 months time
if those halves of the teams have fewer defects than the other.

Very few industrial or research groups have access
to the kinds of resources using by Bente et al. (evidence: in the size years since the
publication of that work, we know of only one of two similar studies). Hence, it
we wish to verify HOW's plans, then some other approach is required that is less
resource intensive.

We propose using prediction systems (built by data mining) as a verification
tool for planners. In domains with data, from which data miners can build
predictors of adequate performance, then those predictors can assess the value
of the changes proposed by HOW.
As shown in \fig{work}, 
project data is divided into two disjoint sets {\em train} and {\em test}
(so {\em train} $\cup$ {\em test} $=\;\emptyset$).
Next, from the train set, we build both a {\em planner} as well
as a {\em predictor}. Our general framework does not   commit to any particular  choice
of {\em planner} or {\em predictor}. For the purposes of this discussion 
our {\em planner}
will be HOW (discussed below)and our
our {\em predictors} will be the Random Forest and Random Forest
Regressor\footnote{Decision tree learners find the attribute whose value's
select for for sections of the data with least variation in the class
variable. Decision tree learners then recurse on each selection~\cite{breiman84}. A random forest is a set of decision trees, each built
from some random subset of the rows and columns. The conclusions
of the forest are a ``vote'' generated by all the trees~\cite{Breiman2001}. 
A standard Random Forest predicts for a discrete class variable while
a Random Forest Regressor generates numeric predictions.} taken from the SciKit
Learn toolkit~\cite{Pedregosa2012}.

Turning now to the {\em test} data, this is based to the {\em predictor}
to generate a baseline {\em before} performance score.
Note that if our {\em predictors} fail to perform adequately on the test data,
then we cannot trust them to comment on our plans. Accordingly,
if that performance is unsatisfactory, we abort.
Else, we (1)~apply the {\em planner} to alter the {\em test} data;
then (2)~apply the {\em predictor} to the altered data {\em test'};
then (3)~return the performance delta $\Delta$.

The clear advantage of this approach is that it can be readily and automatically
applied to domains with historical data. That said, the drawback with this approach
is that our assessment of the plans is only as good as the predictor. To 
mitigate for this {\em problem of poor planners}, it is wise to demand high standards form the {\em
satisfactory} threshold used in this \fig{frame}.

\begin{figure}[!t]
\small
\[  
\begin{array}{r}
project\\
data
\end{array} 
\left\{\begin{array}{l}\mathit{train}
        \left\{\begin{array}{l}
                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{e.g.\;Random\;Forest}\\
                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{e.g.\; HOW}
              \end{array}\right.
       \\
      ~\\
\mathit{test}  
    \left\{\begin{array}{l@{~}l}
           \mathit{before}& =\mathrm{predictor}(\mathit{test})\\
           \mathrm{\bf if\;}\mathit{before} & <  \mathit{satisfactory}\\
           \mathrm{\bf then}  & \mathrm{{\bf return}\; 0}\\~\\
           \mathrm{\bf else} &
           \left\{
            \begin{array}{l}
                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
                \mathrm{{\bf return}\;} \mathit{after} /  \mathit{before}
            \end{array}
          \right.
   \end{array}\right.
\end{array} \right. 
\]

\caption{How to use {\em predictors} for {\em planning}.}\label{fig:work}
\end{figure}

This study explores data sets from two types of domains:
\bi
\item
In one type, we 
build plans to change static code attributes in order to reduce defects in
Java classes.  Here,,
performance can be measured in terms of recall (percent of faulty classes in
the test data detected
by the {\em predictor}). Using engineering judgement, we define satisfactory recall as anything less than 70\%. 
\item 
In the other type, we   build
plans to change configurations settings within a Makefile such that the resulting compiled
code will run faster. Here, performance can be measured in terms of difference
between the predicted runtime $p$ of test case items and their actual runtimes $a$
using  $s= 1- \frac{abs(a - p)}{a}$.
Using engineering judgement, we define satisfactory $s$ as anything less than 70\%. 
\ei
This paper explores 12 static code data sets
Ant, Camel, Ivy, Jedit, JBeans, Log4, Lucene, Poi, Synapse, Velocity, Xalan, Xerces\footnote{Available from the object-oriented defects section of the PROMISE respository openscience.us/repo/defect/ck.} and five configuration data sets Berkeley DB, Apache, SQLite, LLVM,
  x264\footnote{Available from the performance prediction section of PROMISE
  http://openscience.us/repo/performance-predict.}. For each of these data sets,
  we generated 25 {\em before} scores from
  random split train:test splits in the ratio 4:1 train.
 The median {\em before} recalls from  the static code data sets was:
%issue24
\[\mathrm{recall}=\{98,88,86,84,82,81,79,78,78,\overbrace{55,23,20}^{\mathrm{JBeans,Poi,Xalan}}\}\]
The median {\em before} $e$  scores from  the configuration data was
%issue 24
\[s=\{98,97,95,88,92\}\]
Note that we need to abort planning in only $3/17$ of our data sets.

This paper also explores five configuration data sets Berkeley DB, Apache, SQLite, LLVM,
  x264. The median {\em before} scores from these data sets (generated in 
-- which means we cannot generate plans for $3/12$ of our data sets. However, as shown
below, we were able to generate useful plans for the other $9/12$ of the data.

In the other type,

then studied to generate a prediction
on the test data. If the performance on that test data is not
satisfactory, thempredictions:
a {\em before} prediction (on the raw test data); and  an {\em after}
prediction (that is  applying the policies learned from the training data).

The improvement delta $\Delta$ is then computed by dividing {\em after} and {\em before}

This paper presents a novel policy generator called ``HOW''.
HOW is a near-linear time  instance-based
local search algorithm that looks for improvement trends in the data,
then adjusting examples to nudge them along that trend.
The experiments of this paper apply HOW to learning policies for 
software defect reduction and for compiling software that runs faster.
At least for the data explored here, HOW's reductions in defects and runtimes
are very large-- up to 50\% fewer defects and up to 50\% faster runtimes.

The rest of this paper is structured as follows.

@@

\begin{figure}[!b]
\begin{python}[left]
def fastmap(data): 
  """Project data on a line 
     between 2 dist points
  """
  z          = random.choose(data)
  east       = furthest(z, data)
  west       = furthest(east, data)
  data.poles = (west,east)
  c          = dist(west,east)     
  for one in data.members: 
    one.pos = project(west,east,c,one)
  data = sorted(data) # sorted by 'pos'
  return split(data)

def project(west, east, AC, x): 
  "Project x onto line east to west"
  AB = dist(x,west)
  BC = dist(x,east)
  #cosine rule
  return (AB$^2$ + AC$^2$ - AB$^2$)/(2*AC)

def furthest(x,data): 
  # what is furthest from x?
  out, max = x,0
  for y in data:
    d = dist(x,y)
    if d > max: out, max = y, d
  return out

def split(data): # Split at median
   mid = len(data)/2; 
  return data[mid:], data[:mid]

def WHERE(data, scores={}, 
          lvl=10000, prune=True):  
  "Recursively split data."
  if lvl < 1: 
     return data # stop if out of levels
  leafs      = [] # Empty Set
  left,right = fastmap(data)
  west, east = data.poles
  $\omega=\sqrt{\mu}$ # enough data for recursion
  goWest = len(left) > $\omega$  
  goEast = len(right) > $\omega$ 
  if prune: 
    if goEast and better(west,east,scores): 
       goEast = False 
    if goWest and better(east,west,scores): 
       goWest = False 
  if goWest:  
     leafs += where(left,  lvl - 1, prune)  
  if goEast:  
     leafs += where(right, lvl - 1, prune) 
  return leafs

\end{python}
\caption{Clustering with WHERE}
\label{fig:fastmapCode}   
\end{figure}

\begin{figure}[!t]
\begin{python}[right]
def envied(leaves):
  """
  Returns a single exemplary row from 
  a cluster.
  """

def fWeight(train, prune = 0):
  """
  Find the feature weights & prune features
  """

def getContours(train):
  """
  Identify the pairs of clusters
  that can form a contour
  """
  contours = []
  # Clustering with where
  clusters = WHERE(train) 
  near = lambda x: knn(x, clusters)[1]
  for c in clusters:
    one = envied(c)
    two = envied(near(one, clusters))
    contours += [(one, two)]
  return contours

def project(a, b, test):
  if score(a) > score(b):
    good, bad = a,b 
  else:
    good, bad = b,a
  AB = dist(bad,test)
  BC = dist(good,test)
  AC = dist(good,bad)
  # Cosine rule
  x = (AB$^2$ + AC$^2$ - BC$^2$)/(2*AC)
  return $\sqrt{AB^2 - x^2}$

def mutate(me, others, weights = None):
  "Mutate 'me' towards 'others'"
  if weights:
    return [my + ext * f * (good - my) for 
    f, my, good in zip(wights, me, others)]
  else:
    return [my + ext * (good - my) 
    	 for my, good in zip(me, others)]

def nearestContour(contours, instance):
  "Find the nearest contour to an instance"
  return sorted(
        contours,
        key=lambda F: project(
        F[0], F[1], instance), 
        reverse=True)[0]

def HERE():
  contours = getContours()
  weights = fWeight(train)
  for rows in defectiveRows:
    vtx = nearestContour(contours, rows)
    [good, bad] = sorted(vtx
            , key=lambda F: score(F))
    new += mutate(rows, good, weights)
  return new
\end{python}
\caption{Instance-based Planning using HERE}
\label{fig:HERE}   
\end{figure}

@@

\begin{figure*}[htbp!]
  \renewcommand{\baselinestretch}{0.8}\begin{center}
    {\scriptsize
      \begin{tabular}{c|l|p{4.7in}}
        amc & average method complexity & e.g. number of JAVA byte codes\\\hline
        avg\, cc & average McCabe & average McCabe's cyclomatic complexity seen
        in class\\\hline
        ca & afferent couplings & how many other classes use the specific
        class. \\\hline
class. \\\hline
        cam & cohesion amongst classes & summation of number of different
        types of method parameters in every method divided by a multiplication
        of number of different method parameter types in whole class and
        number of methods. \\\hline
        cbm &coupling between methods &  total number of new/redefined methods
        to which all the inherited methods are coupled\\\hline
        cbo & coupling between objects & increased when the methods of one
        class access services of another.\\\hline
        ce & efferent couplings & how many other classes is used by the
        specific class. \\\hline
        dam & data access & ratio of the number of private (protected)
        attributes to the total number of attributes\\\hline
        dit & depth of inheritance tree &\\\hline
        ic & inheritance coupling &  number of parent classes to which a given
        class is coupled (includes counts of methods and variables inherited)
        \\\hline
        lcom & lack of cohesion in methods &number of pairs of methods that do
        not share a reference to an instance variable.\\\hline
        locm3 & another lack of cohesion measure & if $m,a$ are  the number of
        $methods,attributes$
        in a class number and $\mu(a)$  is the number of methods accessing an
        attribute, 
        then
        $lcom3=((\frac{1}{a} \sum, j^a \mu(a, j)) - m)/ (1-m)$.
        \\\hline
        loc & lines of code &\\\hline
        max\, cc & maximum McCabe & maximum McCabe's cyclomatic complexity seen
        in class\\\hline
        mfa & functional abstraction & number of methods inherited by a class
        plus number of methods accessible by member methods of the
        class\\\hline
        moa &  aggregation &  count of the number of data declarations (class
        fields) whose types are user defined classes\\\hline
        noc &  number of children &\\\hline
        npm & number of public methods & \\\hline
        rfc & response for a class &number of  methods invoked in response to
        a message to the object.\\\hline
        wmc & weighted methods per class &\\\hline
        \rowcolor{lightgray}
        defect & defect & Boolean: where defects found in post-release bug-tracking systems.
      \end{tabular}
    }
  \end{center}
  \caption{OO measures used in our defect data sets.  Last line is
    the dependent attribute (whether a defect is reported to  a
    post-release bug-tracking system).}\label{fig:ck}
\end{figure*}

\begin{figure}[!tb]
  \renewcommand{\baselinestretch}{0.8}\begin{center}
{\scriptsize
\begin{tabular}{l@{~~~}l@{~~~}l@{~~~}l@{~~~}l@{~~~}l@{~~~}l@{~~}}
  \hline
  \rowcolor{lightgray}
  Data & Symbol & Training & Testing & Training Samples& Defective &\% Defective \\\hline

Ant & ant & 1.5, 1.6  &1.7 & 644&124&19.25\\

Camel & cam & 1.2, 1.4 & 1.6 & 1480&361 & 24.39\\

Ivy & ivy & 1.1, 1.4 & 2.0  & 352 & 79 & 22.44\\

Jedit & jed & 4.1, 4.2 & 4.3 & 679 & 127 & 18.70\\

Log4j & log & 1.0, 1.1 & 1.2 & 244 & 71 & 29.09\\

Lucene & luc & 2.0, 2.2 & 2.4 & 442 & 235 & 53.16\\

Poi & poi & 2.0, 2.5 & 3.0 & 699 & 285 & 40.77\\

Synapse & syn & 1.0, 1.1 & 1.2 & 379 & 76 & 20.05\\

Velocity & vel & 1.4, 1.5 & 1.6 & 410& 289 & 70.48\\

Xalan & xal &2.5, 2.6 &2.7 & 1688 & 798 & 47.27\\\hline
\end{tabular}}
\end{center}
\caption{Attributes of the defect datasets}\label{fig:ck}
\end{figure}

\begin{figure}[!hbp]
  \renewcommand{\baselinestretch}{1}\begin{center}
{\scriptsize
\begin{tabular}{llllll}
  \hline
  \rowcolor{lightgray}
Project & Domain & Lang. & LOC & Features & Config\\\hline

Berkeley DB CE & Database & C & 219,811 & 18 & 2560\\

Berkeley DB JE & Database & Java & 42,596 & 32  & 400\\

Apache & Web Server & C & 230,277 & 9 & 192\\

SQLite & Database & C & 312,625 & 39 & 3,932,160\textsuperscript{*}\\

LLVM & Compiler & C++ & 47,549 & 11 & 1024\\

x264 & Video Enc. & C& 45,743 & 16 & 1152\\\hline
\end{tabular}}\par\medskip

{ \footnotesize $^*$ Dataset contains 4, 553 configurations for prediction modeling and 100 additional random configurations for prediction evaluation, see \cite{vapp}.}
\end{center}

\caption{Attributes of the Performance Prediction data sets}\label{fig:cpm}
\end{figure}

\begin{figure}[tbp!]
% \subfloat[\label{oracle}]{\includegraphics[width=\linewidth]{{_figs//Oracle}}\\
\includegraphics[width=\linewidth]{_figs/WHAT}
\caption{Experimental Rig}
\label{fig:rig}

%   \subfloat[Experimental Rig for Defect Prediction]{  \includegraphics[width=\linewidth]{_figs/HERE}
%   \label{fig:rig2}}

\end{figure}

%involves two key components. Firstly, there is the oracle (Random Forest with SMOTE) which is used to make predictions. Then we have the 
\section{Data sets}

@@

\subsection{Performance Assessment}

In order to assess the performance of the planner for the defect data set, we use the Cliffs Delta score to measure the probability that the number of bugs in test data before applying the planner is larger than after doing so. In other words, we use the delta score to measure the ability of the planner to effective reduce the number of defects. Given the untreated test instance labeled \textit{Before} of length \textit{M} and the treated instances labeled \textit{After}, the delta score is obtained as follows:
\begin{equation}
\delta = \frac{\#(Before<After) - \#(Before>After)}{M^2}
\label{eq:cliffs}
\end{equation}

In the context of our application, the $\delta$ attains a value of -1 if the number of defects after applying the treatments has reduced to zero. On the other hand $\delta$ will be 0 if there is no change at all. Thus, most of the values reported will lie between 0 and -1. It needs to be highlighted that although $\delta$ can lie between -1 and +1, the in our application the range is limited to -1 and 0. And this is because we only apply the treatments to defective cases.

For the performance prediction data set on the other hand, Cliff's delta measure is not directly applicable. We therefore assess the performance by measuring the gain, given by $\frac{Before}{After}$. The gain takes a value larger than 1 if the run times have been reduced, a value of 1 if there is no change, and a value less than 1 if run times have increased after applying the recommended changes.

In addition to the above, we rank the different various of the planning scheme to identify the best approach. We make use of the Scott-Knott procedure, recommended by Mittas \& Angelis in their 2013 IEEE TSE paper~\cite{sk}, to compute the ranks. It works as follows: A list of treatments \textit{l} is sorted by the Median score. The list \textit{l} is then split into sub-lists m, n in order to maximize the expected value of the differences in the observed performance before and after division. A statistical hypothesis test \textit{H} is applied on the splits \textit{m, n} to check if they are statistically different. If so, Skott-Knott then recurses on each division.

The research conducted by Shepperd and MacDonell~\cite{shepperd12a}, Kampenes~\cite{kampenes07} and Kocaguenli et al.~\cite{}, highlighted that an ``effect size'' in lieu of a mere hypothesis test is required in order to verfiy if two populations are ``significly'' different. An ICSE'11 paper by Arcuri~\cite{} endorsed the use of Vargha and Delaney's A12 effect size for reporting results in software engineering. Thus, for hypothesis testing H in Skott-Knott, we use the A12 test and a non-parametric bootstrap sampling~\cite{}.

\section{Experimental Results}

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=\linewidth]{_figs/histograms.pdf}
\caption{Comparison of defects before and after planning.}
\label{fig:bugs}
\end{figure}
With this in mind, it is easy to note that lower cliffs delta score translates to better the performance.

@@

\begin{figure*}[htbp!]

\noindent\begin{minipage}{0.5\textwidth}
  {\scriptsize \textbf{Ant}\\}
  {\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~}c}}
      \arrayrulecolor{darkgray}
      \rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 & 0.75 &    -0.63  &  0.08 & \quart{2}{7}{5}{144} \\
  1 & 0.75, weighting &    -0.6  &  0.2 & \quart{0}{16}{8}{144} \\
\hline  2 & 0.5, weighting &    -0.42  &  0.11 & \quart{17}{10}{23}{144} \\
  2 & 0.5 &    -0.42  &  0.11 & \quart{19}{9}{23}{144} \\
  2 & 0.75, weighting, Prune=50\% &    -0.4  &  0.11 & \quart{20}{9}{25}{144} \\
\hline  3 & 0.25 &    -0.28  &  0.09 & \quart{31}{7}{35}{144} \\
  3 & 0.25, weighting &    -0.28  &  0.08 & \quart{31}{7}{35}{144} \\
\hline  4 & 0.5, weighting, Prune=50\% &    -0.25  &  0.06 & \quart{35}{5}{38}{144} \\
\hline  5 & 0.25, weighting, Prune=50\% &    -0.19  &  0.05 & \quart{41}{4}{43}{144} \\
\hline  6 & Baseline &    -0.11  &  0.03 & \quart{47}{2}{49}{144} \\
\hline \end{tabular}}
\end{minipage}
\noindent\begin{minipage}{0.5\textwidth}
  \flushleft
  {\scriptsize \textbf{Camel}\\}
{\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~}c}}
\arrayrulecolor{darkgray}
\rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
  1 &  0.75, weighting, Prune=50\% &    -0.41  &  0.1 & \quart{2}{14}{9}{205} \\
  1 &  0.75 &    -0.41  &  0.23 & \quart{1}{32}{9}{205} \\
  1 &  0.75, weighting &    -0.39  &  0.16 & \quart{0}{22}{12}{205} \\
\hline  2 &  0.5, weighting, Prune=50\% &    -0.27  &  0.09 & \quart{19}{12}{29}{205} \\
  2 &  0.5 &    -0.26  &  0.12 & \quart{20}{17}{30}{205} \\
  2 &  0.5, weighting &    -0.24  &  0.09 & \quart{24}{13}{33}{205} \\
\hline  3 &  0.25, weighting, Prune=50\% &    -0.2  &  0.05 & \quart{36}{7}{38}{205} \\
  3 &  0.25, weighting &    -0.19  &  0.07 & \quart{36}{9}{40}{205} \\
  3 &  0.25 &    -0.17  &  0.06 & \quart{37}{8}{43}{205} \\
\hline  4 &  Baseline &    -0.15  &  0.06 & \quart{41}{8}{45}{205} \\
\hline \end{tabular}}
\end{minipage}

>>>>

In should be noted  that the standard 10-way cross-validation
procedure used  in data mining does not necessarily demonstrate G-causality.  In a N-way cross-val experiments, the data is divided into $N$ bins.  
$N-1$ bins are used for learning and the results are tested
on the  remaining bin.
This is repeated $N$ times, each time using a different bin for the test set.
This approach can not be used to demonstrate causality since, if the initial data is collected
over (say) $N$ months, then it is possible that the test set contains observations collected
{\em before}   items seen in the training set.

Jiang et al.~\cite{me11f} and Lumpe et al.~\cite{me11f} propose a modification to cross-validation.
which lets a researcher
demonstrate G-causal defect effects in software engineering.
In this modification, we  divide project data  into $N$ bins, each of with contain data collected at similar t

imes.
They then sort those bins (on the collection date) and use bins $i,j$ to build models that
are applied to bin  $k$ where 
$i<j<k$. Note that this procedure ensures that we are building theories on data seen
before the test data.

We adapt the work of Jiang,Lumpe et al.  as follows.

blah blah data. algorithms. statistics

@@

@@

\section{Data sets}

@@

section{Experimental Results}

@@

With this in mind, it is easy to note that lower cliffs delta score translates to better the performance.

@@
